A user-friendly tool to evaluate the effectiveness of no-take marine reserves

### 2017-12-21 Manuscript.Rmd
Updated manuscript formating with LaTeX template

The first output is a color–coded scorecard intended to provide a general overview of the effectiveness of the reserve. The scorecard provides a global score for the reserve, a general score for each category of indicators, and an individual score for each indicator. The global and category–level scores are determined by the percentage of positive indicators, overall and for each category, respectively. For numeric biological indicators (all but B5), the color is defined by the sign of the interaction term coefficient ($\beta_2$) in eq. 1. For socioeconomic indicators, colors are assigned based on the direction of the slope ($\beta_1$) in eq. 3. Red, yellow, and green are used for $\beta_i<0$, $\beta_i = 0$, and $\beta_i>0$, respectively. The intensity of the color is defined by the significance of the coefficient, testing the null hypothesis of no change (*i.e.* $H_0: \beta_i = 0$) with a Student’s t-test. Cutoff values are $p < 0.05$ and $p < 0.1$. Thus, even in a case where $\beta_i > 0$, if the coefficient is not significant by standard measures (*i.e. * $p>0.1$), the indicator will be assigned a yellow color. A legend (Fig. 2) is provided within the scorecard to aid in the interpretation of these results. Governance indicators are represented simply by red or green. The color is defined based on what literature shows to be a negative (red) or positive (green) factor for a reserve (S3 Table). For example, if the perceived degree of illegal fishing is high, this indicator will be assigned a red color. However, due to the nature of some governance indicators, which require the user to provide a narrative, only some indicators are presented in the scorecard (although all are included in the technical report).

@@

**Fig 3. General location of Isla Natividad (left) and map of the island (right).** The marine reserve polygon is indicated in red, and the approximate location of control sites is indicated by blue squares (B = Babencho, D = La Dulce). Shapefiles for Mexican coastline and the United States were obtained from INEGI @[INEGI] and the tmap R package [@tmap_2017], respectively.

>>>>

The first output is a color–coded scorecard intended to provide a general overview of the effectiveness of the reserve. The scorecard provides a global score for the reserve, a general score for each category of indicators, and an individual score for each indicator. The global and category–level scores are determined by the percentage of positive indicators, overall and for each category, respectively. For numeric biological indicators, the color is defined by the sign of the interaction term coefficient ($\beta_2$) in eq. 1. For socioeconomic indicators, colors are assigned based on the direction of the slope ($\beta_1$) in eq. 3. Red, yellow, and green are used for $\beta_i<0$, $\beta_i = 0$, and $\beta_i>0$, respectively. The intensity of the color is defined by the significance of the coefficient, testing the null hypothesis of no change (*i.e.* $H_0: \beta_i = 0$) with a Student’s t-test. Cutoff values are $p < 0.05$ and $p < 0.1$. Thus, even in a case where $\beta_i > 0$, if the coefficient is not significant by standard measures (*i.e. * $p>0.1$), the indicator will be assigned a yellow color. A legend (Fig. 2) is provided within the scorecard to aid in the interpretation of these results. Governance indicators are represented simply by red or green. The color is defined based on what literature shows to be a negative (red) or positive (green) factor for a reserve (S3 Table). For example, if the perceived degree of illegal fishing is high, this indicator will be assigned a red color. However, due to the nature of some governance indicators, which require the user to provide a narrative, only some indicators are presented in the scorecard (although all are included in the technical report).

@@

**Fig 3. General location of Isla Natividad (left) and map of the island (right).** The marine reserve polygon is indicated in red, and the approximate location of control sites is indicated by blue squares (B = Babencho, D = La Dulce). Shapefiles for Mexican coastline and the United States were obtained from INEGI [@INEGI] and the tmap R package [@tmap_2017], respectively.

### 2017-12-21 Manuscript.tex
Updated manuscript formating with LaTeX template

This sampling design for biological data allows us to use causal
inference techniques {[}21,47{]} to evaluate the effect of the reserve
on biological indicators. The hypothesis that the indicators will
respond to implementation of the reserve is tested by analyzing spatial
and temporal changes in each numeric biological indicator (all but B5)
using generalized linear models {[}21{]}. To account for variations in
the environment and survey conditions, covariates that are gathered
during the underwater ecological surveys are included in the
difference-in-differences model with form:

@@

The first output is a color--coded scorecard intended to provide a
general overview of the effectiveness of the reserve. The scorecard
provides a global score for the reserve, a general score for each
category of indicators, and an individual score for each indicator. The
global and category--level scores are determined by the percentage of
positive indicators, overall and for each category, respectively. For
numeric biological indicators (all but B5), the color is defined by the
sign of the interaction term coefficient (\(\beta_2\)) in eq. 1. For
socioeconomic indicators, colors are assigned based on the direction of
the slope (\(\beta_1\)) in eq. 3. Red, yellow, and green are used for
\(\beta_i<0\), \(\beta_i = 0\), and \(\beta_i>0\), respectively. The
intensity of the color is defined by the significance of the
coefficient, testing the null hypothesis of no change (\emph{i.e.}
\(H_0: \beta_i = 0\)) with a Student's t-test. Cutoff values are
\(p < 0.05\) and \(p < 0.1\). Thus, even in a case where
\(\beta_i > 0\), if the coefficient is not significant by standard
measures (\emph{i.e. } \(p>0.1\)), the indicator will be assigned a
yellow color. A legend (Fig. 2) is provided within the scorecard to aid
in the interpretation of these results. Governance indicators are
represented simply by red or green. The color is defined based on what
literature shows to be a negative (red) or positive (green) factor for a
reserve (S3 Table). For example, if the perceived degree of illegal
fishing is high, this indicator will be assigned a red color. However,
due to the nature of some governance indicators, which require the user
to provide a narrative, only some indicators are presented in the
scorecard (although all are included in the technical report).

@@

\textbf{Fig 3. General location of Isla Natividad (left) and map of the
island (right).} The marine reserve polygon is indicated in red, and the
approximate location of control sites is indicated by blue squares (B =
Babencho, D = La Dulce). Shapefiles for Mexican coastline and the United
States were obtained from the Instituto Nacional de Estadistica,
Geografia e Informatica of Mexico (INEGI:
www.inegi.org.mx/geo/contenidos/geoestadistica/m\_geoestadistico.aspx)
and the tmap R package {[}58{]}, respectively.

In 2006, the Isla Natividad community established a biological baseline
following the data collection protocol described in this study. The
community then implemented two community-based marine reserves within
their TURF {[}8,41,59{]} after establishing a baseline for the
soon-to-be reserves and control sites. Evidence suggest that these
reserves have been effective at enhancing resilience to climate
variations {[}8{]} and preserving genetic diversity of high value
commercial species such as abalone {[}10{]}. These ecological benefits
have been translated into economic benefits, enhancing population
persistence and bolstering abalone fisheries {[}43{]}. For the purpose
of this evaluation, we focused on the ``La Plana / Las Cuevas'' marine
reserve, located at the southern end of the island (Fig. 3) and its
corresponding control site ``La Dulce / Babencho''.

@@

Local fishers (who were trained in scientific diving by the CSO
Comunidad y Biodiversidad, A.C. (COBI; \url{www.cobi.org}), ReefCheck
California, and Stanford University) and personnel from these
institutions performed SCUBA dives to record fish and invertebrate
richness and abundances, as well as fish total length. They recorded
information along 30 m transects, with a sampling window of 2 m x 2 m
following a standardized ReefCheck protocol {[}60{]}. Ecological surveys
were performed yearly in each reserve and corresponding control site(s),
before and after the implementation of the reserve, providing the
requisite time series data inside the reserve and for a suitable control
site. Annual surveys (2006--2016) were carried out in late July -- early
August, performing a total of 242 and 245 transects in the reserve site
for fish and invertebrate surveys, respectively. Similar sampling effort
was applied to the control site, with 221 fish and 222 invertebrate
transects. Between 12 and 27 transects were performed in each site every
year.

@@

Recall that the governance objectives are evaluated based on the
institutions present, not on a specific quantitative linkage between
governance and biological or economic outcomes. Data for this reserve
suggest that the community is strongly organized, which is a likely
driver of the successes reported above {[}61{]}. The first point of
success is the existence of a fishing cooperative that is also
affiliated with a regional federation of cooperatives. These polycentric
governance structures allow various levels of organization that have
been shown to foster communication and cooperation {[}50,57{]};
federations also provide bargain power with governments {[}50,62{]}.
Access to fishing resources is managed through a TURF, permits, and
fishing quotas (for some species). McCay {[}56{]} suggests that the TURF
promotes a sense of stewardship of their resources and incentivizes
sustainable management. Together, these structures enabled a
participative, bottom--up process during the reserve design phase;
opinions of all fishing members --and often non-fishing community
members-- were included. Participation of community members in reserve
surveillance and yearly monitoring indicate commitment and interest, and
allow informal communication of results to uninvolved community members.
Furthermore, the reserve is partially isolated from poaching activity,
and fishers have internal regulations pertaining to the reserves. The
low level of illegal fishing by members of the community and outsiders
both inside and outside the reserve represents another indication of
effectiveness. Governance indicators are summarized in Table 5.

@@

To the best of our knowledge, MAREA is the first tool designed to
evaluate marine reserves. Previous work {[}23,25{]} addressed MPA
evaluation and provided the foundation for our contribution. However,
these did not intended to create user-friendly tools to aid in the
evaluation. As with other conservation management tools, development of
tools that automatize complex calculations can have an important impact
in management {[}63{]}. The use of open data science enables the
creation of open-access tools that can address technical gaps and
inprove management {[}28{]}.

The effectiveness of marine reserves continues to be a matter of debate
{[}11,44,64{]}. With current targets set to increase ocean protection,
it is important that we understand the effects of our interventions
{[}38{]} so we can better inform management {[}47{]}. It is therefore
important that academics, managers, fishers, and CSOs have access to
open access tools like MAREA. This is particularly relevant for Mexico
and other Latin American countries, where management agencies are often
understaffed and underfunded {[}65{]}, or where materials are often not
available in their language. In this context, MAREA provides a simple
and replicable way to align management objectives with performance
indicators. The proposed methodologies, especially the way in which
biological indicators are evaluated, provide valuable information for
managers. We acknowledge there is room for improvement in the way in
which socioeconomic and governance data are analyzed. Despite this,
providing a unifying platform where all indicators can be analyzed and
comprehensively presented represents a valuable step towards effective
evidence---based management {[}47{]}.

@@

\hypertarget{ref-chavez_2003-_m}{}
16. Chavez FP, Ryan J, Lluch-Cota SE, Niquen C M. From anchovies to
sardines and back: Multidecadal change in the pacific ocean. Science.
2003;299: 217--221.
doi:\href{https://doi.org/10.1126/science.1075880}{10.1126/science.1075880}

@@

\hypertarget{ref-sandwich_2014}{}
49. Zeileis A. Econometric computing with hc and hac covariance matrix
estimators. Journal of Statistical Software. 2004;11: 1--17.

@@

\hypertarget{ref-tmap_2017}{}
58. Tennekes M. Tmap: Thematic maps {[}Internet{]}. 2017. Available:
\url{https://CRAN.R-project.org/package=tmap}

\hypertarget{ref-afflerbach_2014-HP}{}
59. Afflerbach JC, Lester SE, Dougherty DT, Poon SE. A global survey of
-reserves, territorial use rights for fisheries coupled with marine
reserves. Global Ecology and Conservation. 2014;2: 97--106.
doi:\href{https://doi.org/10.1016/j.gecco.2014.08.001}{10.1016/j.gecco.2014.08.001}

\hypertarget{ref-suman_2010-ez}{}
60. Suman CS, Saenz-Arroyo A, Dawson C, Luna MC. Manual de instruccion
de reef check california: Guia de instruccion para el monitoreo del
bosque de sargazo en la peninsula de baja california. Pacific Palisades,
CA, USA: Reef Check Foundation; 2010.

\hypertarget{ref-gutirrez_2011-0U}{}
61. Gutiérrez NL, Hilborn R, Defeo O. Leadership, social capital and
incentives promote successful fisheries. Nature. 2011;470: 386--389.
doi:\href{https://doi.org/10.1038/nature09689}{10.1038/nature09689}

\hypertarget{ref-finkbeiner_2015-87}{}
62. Finkbeiner EM, Basurto X. Re-defining co-management to facilitate
small-scale fisheries reform: An illustration from northwest mexico.
Marine Policy. 2015;51: 433--441.
doi:\href{https://doi.org/10.1016/j.marpol.2014.10.010}{10.1016/j.marpol.2014.10.010}

\hypertarget{ref-ball_2009-qi}{}
63. Ball IR, Possingham HP, Watts ME. Marxan and relatives: Software for
spatial conservation prioritization. Spatial conservation prioritization
quantitative methods \& computational tools. United Kingdom: Oxford
University Press; 2009. pp. 185--195. Available:
\url{https://espace.library.uq.edu.au/view/UQ:200259}

\hypertarget{ref-padleton_2017-vn}{}
64. Padleton L, Aghmadia G, Browman H, Thurstand R, Kaplan D, Bartolino
V. Debating the effectiveness of marine protected areas. ICES Journal of
Marine Science. 2017;
doi:\href{https://doi.org/10.1093/icesjms/fsx154}{10.1093/icesjms/fsx154}

\hypertarget{ref-lundquist_2005-OL}{}
65. Lundquist CJ, Granek EF. Strategies for successful marine
conservation: Integrating socioeconomic, political, and scientific
factors. Conserv Biol. 2005;19: 1771--1778.
doi:\href{https://doi.org/10.1111/j.1523-1739.2005.00279.x}{10.1111/j.1523-1739.2005.00279.x}

>>>>

This sampling design for biological data allows us to use causal
inference techniques {[}21,47{]} to evaluate the effect of the reserve
on biological indicators. The hypothesis that the indicators will
respond to implementation of the reserve is tested by analyzing spatial
and temporal changes in each numeric biological indicator using
generalized linear models {[}21{]}. To account for variations in the
environment and survey conditions, covariates that are gathered during
the underwater ecological surveys are included in the
difference-in-differences model with form:

@@

The first output is a color--coded scorecard intended to provide a
general overview of the effectiveness of the reserve. The scorecard
provides a global score for the reserve, a general score for each
category of indicators, and an individual score for each indicator. The
global and category--level scores are determined by the percentage of
positive indicators, overall and for each category, respectively. For
numeric biological indicators, the color is defined by the sign of the
interaction term coefficient (\(\beta_2\)) in eq. 1. For socioeconomic
indicators, colors are assigned based on the direction of the slope
(\(\beta_1\)) in eq. 3. Red, yellow, and green are used for
\(\beta_i<0\), \(\beta_i = 0\), and \(\beta_i>0\), respectively. The
intensity of the color is defined by the significance of the
coefficient, testing the null hypothesis of no change (\emph{i.e.}
\(H_0: \beta_i = 0\)) with a Student's t-test. Cutoff values are
\(p < 0.05\) and \(p < 0.1\). Thus, even in a case where
\(\beta_i > 0\), if the coefficient is not significant by standard
measures (\emph{i.e. } \(p>0.1\)), the indicator will be assigned a
yellow color. A legend (Fig. 2) is provided within the scorecard to aid
in the interpretation of these results. Governance indicators are
represented simply by red or green. The color is defined based on what
literature shows to be a negative (red) or positive (green) factor for a
reserve (S3 Table). For example, if the perceived degree of illegal
fishing is high, this indicator will be assigned a red color. However,
due to the nature of some governance indicators, which require the user
to provide a narrative, only some indicators are presented in the
scorecard (although all are included in the technical report).

@@

\textbf{Fig 3. General location of Isla Natividad (left) and map of the
island (right).} The marine reserve polygon is indicated in red, and the
approximate location of control sites is indicated by blue squares (B =
Babencho, D = La Dulce). Shapefiles for Mexican coastline and the United
States were obtained from INEGI {[}58{]} and the tmap R package
{[}59{]}, respectively.

In 2006, the Isla Natividad community established a biological baseline
following the data collection protocol described in this study. The
community then implemented two community-based marine reserves within
their TURF {[}8,41,60{]} after establishing a baseline for the
soon-to-be reserves and control sites. Evidence suggest that these
reserves have been effective at enhancing resilience to climate
variations {[}8{]} and preserving genetic diversity of high value
commercial species such as abalone {[}10{]}. These ecological benefits
have been translated into economic benefits, enhancing population
persistence and bolstering abalone fisheries {[}43{]}. For the purpose
of this evaluation, we focused on the ``La Plana / Las Cuevas'' marine
reserve, located at the southern end of the island (Fig. 3) and its
corresponding control site ``La Dulce / Babencho''.

@@

Local fishers (who were trained in scientific diving by the CSO
Comunidad y Biodiversidad, A.C. (COBI; \url{www.cobi.org}), ReefCheck
California, and Stanford University) and personnel from these
institutions performed SCUBA dives to record fish and invertebrate
richness and abundances, as well as fish total length. They recorded
information along 30 m transects, with a sampling window of 2 m x 2 m
following a standardized ReefCheck protocol {[}61{]}. Ecological surveys
were performed yearly in each reserve and corresponding control site(s),
before and after the implementation of the reserve, providing the
requisite time series data inside the reserve and for a suitable control
site. Annual surveys (2006--2016) were carried out in late July -- early
August, performing a total of 242 and 245 transects in the reserve site
for fish and invertebrate surveys, respectively. Similar sampling effort
was applied to the control site, with 221 fish and 222 invertebrate
transects. Between 12 and 27 transects were performed in each site every
year.

@@

Recall that the governance objectives are evaluated based on the
institutions present, not on a specific quantitative linkage between
governance and biological or economic outcomes. Data for this reserve
suggest that the community is strongly organized, which is a likely
driver of the successes reported above {[}62{]}. The first point of
success is the existence of a fishing cooperative that is also
affiliated with a regional federation of cooperatives. These polycentric
governance structures allow various levels of organization that have
been shown to foster communication and cooperation {[}50,57{]};
federations also provide bargain power with governments {[}50,63{]}.
Access to fishing resources is managed through a TURF, permits, and
fishing quotas (for some species). McCay {[}56{]} suggests that the TURF
promotes a sense of stewardship of their resources and incentivizes
sustainable management. Together, these structures enabled a
participative, bottom--up process during the reserve design phase;
opinions of all fishing members --and often non-fishing community
members-- were included. Participation of community members in reserve
surveillance and yearly monitoring indicate commitment and interest, and
allow informal communication of results to uninvolved community members.
Furthermore, the reserve is partially isolated from poaching activity,
and fishers have internal regulations pertaining to the reserves. The
low level of illegal fishing by members of the community and outsiders
both inside and outside the reserve represents another indication of
effectiveness. Governance indicators are summarized in Table 5.

@@

To the best of our knowledge, MAREA is the first tool designed to
evaluate marine reserves. Previous work {[}23,25{]} addressed MPA
evaluation and provided the foundation for our contribution. However,
these did not intended to create user-friendly tools to aid in the
evaluation. As with other conservation management tools, development of
tools that automatize complex calculations can have an important impact
in management {[}64{]}. The use of open data science enables the
creation of open-access tools that can address technical gaps and
inprove management {[}28{]}.

The effectiveness of marine reserves continues to be a matter of debate
{[}11,44,65{]}. With current targets set to increase ocean protection,
it is important that we understand the effects of our interventions
{[}38{]} so we can better inform management {[}47{]}. It is therefore
important that academics, managers, fishers, and CSOs have access to
open access tools like MAREA. This is particularly relevant for Mexico
and other Latin American countries, where management agencies are often
understaffed and underfunded {[}66{]}, or where materials are often not
available in their language. In this context, MAREA provides a simple
and replicable way to align management objectives with performance
indicators. The proposed methodologies, especially the way in which
biological indicators are evaluated, provide valuable information for
managers. We acknowledge there is room for improvement in the way in
which socioeconomic and governance data are analyzed. Despite this,
providing a unifying platform where all indicators can be analyzed and
comprehensively presented represents a valuable step towards effective
evidence---based management {[}47{]}.

@@

\hypertarget{ref-chavez_2003-_m}{}
16. Chavez FP. From anchovies to sardines and back: Multidecadal change
in the pacific ocean. Science. 2003;299: 217--221.
doi:\href{https://doi.org/10.1126/science.1075880}{10.1126/science.1075880}

@@

\hypertarget{ref-zeileis_2004-7n}{}
49. Zeileis A. Econometric computing with hc and hac covariance matrix
estimators. J Stat Softw. 2004;11.
doi:\href{https://doi.org/10.18637/jss.v011.i10}{10.18637/jss.v011.i10}

@@

\hypertarget{ref-INEGI}{}
58. Estadistica Geografia e Informatica de Mexico IN de. Marco
geoestadistico nacional. 2017. Available:
\url{www.inegi.org.mx/geo/contenidos/geoestadistica/m_geoestadistico.aspx}

\hypertarget{ref-tmap_2017}{}
59. Tennekes M. Tmap: Thematic maps {[}Internet{]}. 2017. Available:
\url{https://CRAN.R-project.org/package=tmap}

\hypertarget{ref-afflerbach_2014-HP}{}
60. Afflerbach JC, Lester SE, Dougherty DT, Poon SE. A global survey of
-reserves, territorial use rights for fisheries coupled with marine
reserves. Global Ecology and Conservation. 2014;2: 97--106.
doi:\href{https://doi.org/10.1016/j.gecco.2014.08.001}{10.1016/j.gecco.2014.08.001}

\hypertarget{ref-suman_2010-ez}{}
61. Suman CS, Saenz-Arroyo A, Dawson C, Luna MC. Manual de instruccion
de reef check california: Guia de instruccion para el monitoreo del
bosque de sargazo en la peninsula de baja california. Pacific Palisades,
CA, USA: Reef Check Foundation; 2010.

\hypertarget{ref-gutirrez_2011-0U}{}
62. Gutiérrez NL, Hilborn R, Defeo O. Leadership, social capital and
incentives promote successful fisheries. Nature. 2011;470: 386--389.
doi:\href{https://doi.org/10.1038/nature09689}{10.1038/nature09689}

\hypertarget{ref-finkbeiner_2015-87}{}
63. Finkbeiner EM, Basurto X. Re-defining co-management to facilitate
small-scale fisheries reform: An illustration from northwest mexico.
Marine Policy. 2015;51: 433--441.
doi:\href{https://doi.org/10.1016/j.marpol.2014.10.010}{10.1016/j.marpol.2014.10.010}

\hypertarget{ref-ball_2009-qi}{}
64. Ball IR, Possingham HP, Watts ME. Marxan and relatives: Software for
spatial conservation prioritization. Spatial conservation prioritization
quantitative methods \& computational tools. United Kingdom: Oxford
University Press; 2009. pp. 185--195. Available:
\url{https://espace.library.uq.edu.au/view/UQ:200259}

\hypertarget{ref-padleton_2017-vn}{}
65. Padleton L, Aghmadia G, Browman H, Thurstand R, Kaplan D, Bartolino
V. Debating the effectiveness of marine protected areas. ICES Journal of
Marine Science. 2017;
doi:\href{https://doi.org/10.1093/icesjms/fsx154}{10.1093/icesjms/fsx154}

\hypertarget{ref-lundquist_2005-OL}{}
66. Lundquist CJ, Granek EF. Strategies for successful marine
conservation: Integrating socioeconomic, political, and scientific
factors. Conserv Biol. 2005;19: 1771--1778.
doi:\href{https://doi.org/10.1111/j.1523-1739.2005.00279.x}{10.1111/j.1523-1739.2005.00279.x}
